---
title: "magmavizR-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{magmavizR-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Usage

The magmavizR library can be loaded by using the commands below:

```{r echo=TRUE, message=FALSE, warning=FALSE}
library("magmavizR")
penguins_data <- palmerpenguins::penguins
penguins_data
```

The four data visualization functions included in the package along with the usage are outlined below:

### Boxplot

Returns a boxplot based on the data frame, a numerical feature to view the distribution of and a categorical feature to bucket data into categories. Additionally, there is a boolean option to facet the boxplots into separate charts.

```{r echo=TRUE, message=FALSE, warning=FALSE}
boxplot(penguins_data, species, bill_length_mm, facet = TRUE)
```

### Correlation plot

Returns a correlation plot based on the numerical features present in the data frame. Additionally, it will print the correlated numerical feature pairs along with their correlation values.

```{r echo=TRUE, message=FALSE, warning=FALSE}
corrplot(penguins_data, print_corr = TRUE, title = "Correlation")
```

### Histogram

Returns a histogram based on the data frame and a numeric feature to plot on the x-axis. The y-axis will display the result of the following aggregating functions:

-   count

-   ncount

-   density

-   ndensity

-   width

```{r echo=TRUE, message=FALSE, warning=FALSE}
histogram(penguins_data, bill_length_mm, "..count..")
```

### Scatterplot

Returns a scatterplot based on the data frame and two numerical feature names passed as the required inputs. There are auxiliary inputs that provide the flexibility to:

-   Color code or change the shape of the data points on a categorical variable

-   Set a title to the plot, x-axis, y-axis and color legend

-   Change the opacity and size of the data points

-   Set the scale of the x-axis and y-axis to start from zero

```{r echo=TRUE, message=FALSE, warning=FALSE}
scatterplot(penguins_data, bill_length_mm, flipper_length_mm, species, "Bill and Flipper length clusters by Species", 0.5, 2.5, "Bill length (mm)", "Flipper length (mm)", "", FALSE, FALSE, TRUE)
```
